---
title: 'Chapter 3: One-parameter models'
author: 'Shao-Ting Chiu'
date: '20220901'
jupyter: 'r'
---


## Key messages

- One-parameter models
    - Binomial model
    - Poisson model
- Bayesian data analysis
    - Conjugate prior distribution
    - Predictive distribution
    - Confidence regions

## Some one-parameter models


## Bayesian prediction

### The marginal
$$\begin{aligned}
  p(y) &= \int p(y,\theta)d\theta\\
       &= \int_{\Theta} p(y|\theta)p(\theta)d\theta\\
\end{aligned}$$

### Posterior predictive distribution

Let $\bar{Y}$ be a data point that is yet to be observed.

$$\begin{aligned}
  p(\bar{y}|y) &= \int_{\Theta} p(\bar{y}, \theta|y)d\theta\\
               &= \int_{\Theta} p(\bar{y}|\theta,y)p(\theta|y)d\theta\\
\end{aligned}$$


## Jeffreys prior

- [Notes 12. The Jeffreys Prior](https://www2.stat.duke.edu/courses/Fall11/sta114/jeffreys.pdf)


## Gamma Distribution

> Conjuagate prior of Poisson data

$$p(\theta) = \frac{b^a}{\Gamma(a)}\theta^{a-1} e^{-b\theta}I_{0,\infty}(\theta)$$

- posterior of poisson data
$$E(\theta|y) = \frac{a+n\bar{y}}{b+n} = \frac{b}{b+n}\frac{a}{b} + \frac{n}{b+n}\frac{n\bar{y}}{n} = (1-\omega_n)E(\theta) + \omega_n \bar{y}$$

```{r}
a=1; b=1
curve(dgamma(x,a,b),0, 10)
```

```{r}
a=4; b=4
curve(dgamma(x,a,b),0, 10)
```

```{r}
a=16; b=4
curve(dgamma(x,a,b),0, 10)
```


---

## Testing

```{r}
a<-1; b<-1
n1<-111; sy1<-217
n2<-44; sy2<-66
qgamma(c(.025,.975), a+sy1, b+n1)
```

```{r}
plot(c(1,2,3,4,5),c(1,4,9,16,25))
```



## Installation
R installation: https://www.drdataking.com/post/how-to-add-existing-r-to-jupyter-notebook/